V4046\,Sgr high-resolution 1.3\,mm continuum ALMA observations: a disc with a thin ring

### 2020-10-08 Draft_V4046Sgr_mnras.tex
Rename Draft_mnras.tex to Draft_V4046Sgr_mnras.tex

% Abstract of the paper
\begin{abstract}
    %context
    V4046 Sgr is a nearby pre-main sequence spectroscopic binary system that has a gas-rich circumbinary disc.
    %aim
    We report on new $\sim20$\,mas resolution Atacama Large Millimeter Array (ALMA) observations of the radio continuum at 1.3\,mm, and compare with archival SPHERE-IRDIS differential polarised imaging (DPI) of scattered-light at H-band (1.65\,$\micron$) to reach conclusions on the disc structure. 
    %methods
    We use a parametric a radiative transfer model of the circumbinary disc, computed with \textsc{radmc3d} to interpret the available data, including the spectral-energy-distribution, in terms of a physical model. In addition, we explore decompositions of the imaging data in polar coordinates to better constrain the disc radial structure. 
    %results
    We report a two ringed structure for the millimetre-size dust population with a narrow, previously unseen, less than 4.17$\pm$0.94\,au-wide inner ring centred at 13.46$\pm$0.43\,au. This ring is the finest reported so far in a gas rich protoplanetary disc. The ring is surrounded by a $\sim$9\,au-wide gap, flanked by a very sharp outer ring starting at 24\,au. The radial profile of the outer ring exhibits an intriguing discontinuity at $\sim$35\,au. The RT model reproduces accurately the available data. We predict a three ringed structure for the small-dust population with a inner ring at $\sim$6\,au hiding under the coronagraph in the SPHERE observations. The shadow of the secondary reported previously from the scattered-light data is also reproduced by the RT model. 
    %conclusions
    The DPI image reveals strong forward-scattering at 1.65\,$\micron$, which we reproduced with rather large dust grains of no less than 0.4\,$\micron$, hinting at a deficit of very-small-grains. Using the parametric model scale height as an estimated height, we found that the thin inner ring is stable. Deeper ALMA observations may confirm the lopsided 5\,au ring.

@@

Disc structures can be explored by means of comparing mm continuum and scattered light images. These observations allow us to see the distribution of different populations of dust particles, while ALMA traces the millimetre-sized grains settled in the mid-plane of the disc, scattered light accounts for photons reflected off small micron-sized dust at the disc surface layer, high above the mid-plane. By comparing both, is easier to determine whether asymmetries in scattered light correspond to surface ripples or deep structures intrinsic to the underlying density distribution.

@@

For our analysis of the ALMA image we used an image reconstruction strategy, using the non-parametric image synthesis of the \textsc{uvmem} package \citep{2006ApJ...639..951C, 2018A&C....22...16C} it is possible to super-resolve the clean beam, obtaining an effective angular resolution $\sim$2-3 times higher than in the clean image with natural weights. The model obtains this higher resolution through fitting the data using Chi-squared minimisation. Here we adopted a pure $\chi ^2$ model image. The resulting \textsc{uvmem} image shown in Fig. \ref{fig:two} exhibits previously unseen substructure of the disc. This image reveals that the disc features two rings of large dust grains with a broad gap between them. The wide and brighter outer ring has its peak intensity at $\sim$30\,au, has a surface brightness break at $\sim$35\,au, and reaches out to $\sim$60\,au. The thin and weak inner ring had never seen before, but \citet{Ru_z_Rodr_guez_2019} did anticipate its existence.

@@

The scattered light image shown in Fig. \ref{fig:two} reveals the presence of a double ring structure in the micron-sized dust distribution. The observed morphology presents an inner cavity of $\sim$10\,au in radius and two rings located at 14.10$\pm$0.01 and 24.62$\pm$0.08\,au with a small gap between them at $\sim$20\,au. The observed second ring might correspond to the inner wall of the 1.3\,mm continuum emission outer ring \citep{Ru_z_Rodr_guez_2019}. Two other important features that are present in the image are: the brightness asymmetry and the shadows projected on the disc produced by the close binary system as they eclipse each other, described in detail by \citet{dOrazi}.

\citet{dOrazi} also measured the binary phase in the scattered light observation around 106\,$\degr$, setting $\phi=$0 from the North and increasing clockwise. Using this measurement, the binary phase was calculated at the time of the ALMA observation at $\phi=$13\,$\degr$. There is no clear signal of any shadow similar to the ones present in the DPI image, this could be explained due to the slower cooling of larger grains (cite Casassus?***).

@@

The key conclusions of our analysis are as follows.
\begin{itemize}
  \item We report a narrow inner ring for the 1.3\,mm continuum located at 13.46$\pm$0.43\,au from the stars and has an estimated width of 4.17$\pm$0.94\,au. The location of this ring is similar to the inner ring observed in the scattered light image, revealing that the ring includes a considerable mass of millimetre-sized grains of around 3\,M$_{\earth}$. Using the parametric model scale height value ($h= $ 0.28\,au at 13.46\,au) we have that the ring width is $\sim$15 times its estimated height, making it a stable ring.

@@

  \item As our parametric model accounts for the SED of the system, the disc requires the existence of a sub-micron dust population close (<5\,au) to the stars. We also predict the existence of another thin ring at $\sim$6\,au, about 3\,au-wide and made of small dust grains that lies under the coronagraph of the SPHERE images. Additionally, the weak central emission at 1.3\,mm could be part of this ring.

>>>>

% Abstract of the paper
\begin{abstract}
    %context
    V4046 Sgr is a nearby pre-main sequence spectroscopic binary system that has a gas-rich circumbinary disc.
    %aim
    We report on new $\sim$20\,mas resolution Atacama Large Millimeter Array (ALMA) observations of the radio continuum at 1.3\,mm, and compare with archival SPHERE-IRDIS differential polarised imaging (DPI) of scattered-light at H-band (1.65\,$\micron$) to reach conclusions on the disc structure. 
    %methods
    We use a parametric a radiative transfer model of the circumbinary disc, computed with \textsc{radmc3d} to interpret the available data, including the spectral-energy-distribution, in terms of a physical model. In addition, we explore decompositions of the imaging data in polar coordinates to better constrain the disc radial structure. 
    %results
    We report a two ringed structure for the millimetre-size dust population with a narrow, previously unseen, less than 4.17$\pm$0.94\,au-wide inner ring centred at 13.46$\pm$0.43\,au. This ring is the finest reported so far in a gas rich protoplanetary disc. The ring is surrounded by a $\sim$9\,au-wide gap, flanked by a very sharp outer ring starting at 24\,au. The radial profile of the outer ring exhibits an intriguing discontinuity at $\sim$35\,au. The RT model reproduces accurately the available data. We predict a three ringed structure for the small-dust population with a inner ring at $\sim$6\,au hiding under the coronagraph in the SPHERE observations. The shadow of the secondary reported previously from the scattered-light data is also reproduced by the RT model. 
    %conclusions
    The DPI image reveals strong forward-scattering at 1.65\,$\micron$, which we reproduced with rather large dust grains of no less than 0.4\,$\micron$, hinting at a deficit of very-small-grains. Using the parametric model scale height as an estimated height, we found that the thin inner ring is stable. Deeper ALMA observations may confirm the lopsided 5\,au ring.

@@

Disc structures can be explored by means of comparing mm continuum and scattered-light images. These observations allow us to see the distribution of different populations of dust particles, while ALMA traces the millimetre-sized grains settled in the mid-plane of the disc, scattered-light imaging accounts for photons reflected off small micron-sized dust at the disc surface layer, high above the mid-plane. By comparing both, is easier to determine whether asymmetries in scattered-light correspond to surface ripples or deep structures intrinsic to the underlying density distribution.

@@

For our analysis of the ALMA image we used an image reconstruction strategy, using the non-parametric image synthesis of the \textsc{uvmem} package \citep{2006ApJ...639..951C, 2018A&C....22...16C} it is possible to super-resolve the clean beam, obtaining an effective angular resolution $\sim$2-3 times higher than in the clean image with natural weights. The model obtains this higher resolution through fitting the data using Chi-squared minimisation. Here we adopted a pure $\chi ^2$ model image. The resulting \textsc{uvmem} image shown in Fig. \ref{fig:two} exhibits previously unseen substructure of the disc. This image reveals that the disc features two rings of large dust grains with a broad gap between them. The wide and brighter outer ring has its peak intensity at $\sim$30\,au, has a surface brightness break at $\sim$35\,au, and reaches out to $\sim$60\,au. The thin and weak inner ring had never seen before, however \citet{Ru_z_Rodr_guez_2019} did anticipate its existence.

@@

The scattered-light image shown in Fig. \ref{fig:two} reveals the presence of a double ring structure in the micron-sized dust distribution. The observed morphology presents an inner cavity of $\sim$10\,au in radius and two rings located at 14.10$\pm$0.01 and 24.62$\pm$0.08\,au with a small gap between them at $\sim$20\,au. The observed second ring might correspond to the inner wall of the 1.3\,mm continuum emission outer ring \citep{Ru_z_Rodr_guez_2019}. Two other important features that are present in the image are: the brightness asymmetry and the shadows projected on the disc produced by the close binary system as they eclipse each other, described in detail by \citet{dOrazi}.

\citet{dOrazi} also measured the binary phase in the scattered-light observation around 106\,$\degr$, setting $\phi=$0 from the North and increasing clockwise. Using this measurement, the binary phase was calculated at the time of the ALMA observation at $\phi=$13\,$\degr$. There is no clear signal of any shadow similar to the ones present in the DPI image, this could be explained due to the slower cooling of larger grains (cite Casassus?***).

@@

The key conclusions of our analysis are as follows.
\begin{itemize}
  \item We report a narrow inner ring for the 1.3\,mm continuum located at 13.46$\pm$0.43\,au from the stars and has an estimated width of 4.17$\pm$0.94\,au. The location of this ring is similar to the inner ring observed in the scattered-light image, revealing that the ring includes a considerable mass of millimetre-sized grains of around 3\,M$_{\earth}$. Using the parametric model scale height value ($h= $ 0.28\,au at 13.46\,au) we have that the ring width is $\sim$15 times its estimated height, making it a stable ring.

@@

  \item As our parametric model accounts for the SED of the system, the disc requires the existence of a sub-micron dust population close (<5\,au) to the stars. We also predict the existence of another thin ring at $\sim$6\,au, about 3\,au-wide and made of small dust grains that lies under the coronagraph of the scattered-light image. Additionally, the weak central emission at 1.3\,mm could be part of this ring.
